---
title: "LUNDRIGAN_Skills_Test_3"
author: "Jared Lundrigan"
date: "2022-11-22"
output: html_document
---

# Analysis of a 1995 Faculty Salaries Data Set

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading packages

```{r, results= FALSE , message = FALSE}
library(tidyverse)
library(janitor)
library(modelr)
library(broom)
```

### Loading the dataset and cleaning it

```{r message=FALSE, warning=FALSE, results=FALSE}

list.files()
df <- read_csv("FacultySalaries_1995.csv")
df <- clean_names(df)

df1 <- pivot_longer(df, cols = c("num_full_profs","num_assoc_profs", "num_assist_profs"),
             names_to = "profs",
             values_to = "total_profs")
  
df2 <- pivot_longer(df1, cols = c("avg_full_prof_salary", "avg_assoc_prof_salary", "avg_assist_prof_salary"),
               names_to = "prof_type_salary",
               values_to = "salary_by_type")

# changing the names of the values
df2["prof_type_salary"][df2["prof_type_salary"] == "avg_full_prof_salary"] <- "Full"
df2["prof_type_salary"][df2["prof_type_salary"] == "avg_assist_prof_salary"] <- "Assist"
df2["prof_type_salary"][df2["prof_type_salary"] == "avg_assoc_prof_salary"] <- "Assoc"

```

### Graphing the salary by rank and college tier

```{r message=F, warning=F}
df2 %>% 
  filter(tier != "VIIB") %>%   # filtered tier VIIB since we're only looking at I, IIA, IIB
  ggplot(aes(x=prof_type_salary,
             y=salary_by_type,
             fill = prof_type_salary)) +
  geom_boxplot() +
  facet_wrap(~tier) +
  theme_minimal() +
  labs(x = "Rank", y = "Salary", ) +
  theme(axis.text.x = element_text(angle = 45)) +
  scale_fill_discrete(name = "Rank")
```

### An ANOVA model that tests the influence of "State", "Tier", and "Rank" on "Salary

```{r, attr.output='style="max-height: 200px;"'}

df3 <- df2

colnames(df3)[14] = "rank"

colnames(df3)[15] = "salary"

# I had to change from character to factor for the model to work
df3$state <- as.factor(df3$state)
df3$tier <- as.factor(df3$tier)
df3$rank <- as.factor(df3$rank)

mod1 <- glm(data = df3, formula = salary ~ state + tier + rank)

summary(mod1)

```

<br>
<br>

# Analysis of a Juniper Oils Data Set
### Data collected from a collaboration between Young Living, Inc. and UVU Microbiology. It involves data on dead cedar trees and their essential oil content.

### Cleaning the juniper dataset
```{r, results=F, message=F}
list.files()
juniper <- read_csv("Juniper_Oils.csv")
juniper1 <- clean_names(juniper)

colnames(juniper1)

# creating a new column to calculate years since burn
years_since_burn <- 2022 - juniper1$burn_year
juniper2 <- pivot_longer(juniper1, cols = c(11:33), names_to = "chemical_compounds", 
                         values_to = "concentration")

```

### Graphing the juniper data
```{r, message=F, warning=F}

juniper2 %>% 
  ggplot(aes(x = years_since_burn,
             y = concentration)) +
  geom_smooth() +
  facet_wrap(~chemical_compounds, scales = "free") +
  theme_minimal() +
  labs(x = "YearsSinceBurn", y = "Concentration")

```


### Generalized linear model to find which chemicals show statistically significant concentrations affected by years since burn


```{r, attr.output='style="max-height: 200px;"'}

jmod <- glm(data = juniper2, formula = years_since_burn ~ concentration * chemical_compounds)

jmod_tibble <- tidy(jmod) %>% 
  filter(p.value < 0.5)

jmod_tibble
```

<br>
<br>



